\Large Analyzing the Supreme Court Citation Network

### 2017-12-22 Tex/paper.tex
added coefficient plots to ms

\documentclass[headsepline=true, abstracton]{scrartcl}
%\usepackage[ngerman]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage{verbatim}
\usepackage[a4paper,text={140mm,215mm},centering,headsep=10mm,footskip=10mm]{geometry}
\usepackage[urlcolor=black,colorlinks=true,linkcolor=black,citecolor=black,bookmarks]{hyperref}
\usepackage{aliascnt}
\usepackage{lmodern}
\usepackage{mdwlist}
\usepackage{natbib}
\usepackage{listings}
\usepackage[table,xcdraw]{xcolor}
\usepackage[activate]{pdfcprot}
\usepackage{graphicx}
\usepackage{slashed}
\usepackage{mathrsfs}
\usepackage{geometry}
\usepackage{float}
\usepackage{oldgerm}
\usepackage{setspace}
\usepackage{dsfont}
%\usepackage{mathtools}
\usepackage[all]{xy}
\usepackage{cite}
%\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{algorithm2e}
\usepackage{url}
\pagestyle{headings}
\newcommand{\myclearpage}{\clearpage}
\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
\newenvironment{gelaber}{}{}
\newenvironment{preamble}{}{}
\newcommand{\tostar}{\overset{*}{\lower0.5em\hbox{$\smash{\scriptscriptstyle\rightharpoonup}$}}}
\newtheorem{mydef}{Definition}
\newtheorem{bem}{Bemerkung}
\usepackage{tikz}
\newcommand\circlearound[1]{%
  \tikz[baseline]\node[draw,shape=circle,anchor=base] {#1} ;}

@@

We are not the first to study what predicts citations on the Supreme Court...

We are the first to examine at the level of the citation, and, as a result, be able to incorporate complex dependence into our explanation of citations.

\section{Complex Dependence in Court Citations}

@@

>>>>

\documentclass[headsepline=true, abstracton]{scrartcl}
%\usepackage[ngerman]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage{verbatim}
\usepackage[a4paper,text={160mm,215mm},centering,headsep=10mm,footskip=10mm]{geometry}
\usepackage[urlcolor=black,colorlinks=true,linkcolor=black,citecolor=black,bookmarks]{hyperref}
\usepackage{aliascnt}
\usepackage{lmodern}
\usepackage{mdwlist}
\usepackage{natbib}
\usepackage{listings}
\usepackage[table,xcdraw]{xcolor}
\usepackage[activate]{pdfcprot}
\usepackage{graphicx}
\usepackage{slashed}
\usepackage{mathrsfs}
\usepackage{geometry}
\usepackage{float}
\usepackage{oldgerm}
\usepackage{setspace}
\usepackage{dsfont}
%\usepackage{mathtools}
\usepackage[all]{xy}
\usepackage{cite}
%\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{algorithm2e}
\usepackage{url}
\pagestyle{headings}
\newcommand{\myclearpage}{\clearpage}
\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
\newenvironment{gelaber}{}{}
\newenvironment{preamble}{}{}
\newcommand{\tostar}{\overset{*}{\lower0.5em\hbox{$\smash{\scriptscriptstyle\rightharpoonup}$}}}
\newtheorem{mydef}{Definition}
\newtheorem{bem}{Bemerkung}
\usepackage{tikz}
\newcommand\circlearound[1]{%
  \tikz[baseline]\node[draw,shape=circle,anchor=base] {#1} ;}

@@

We are not the first to ask what predicts the citations in US Supreme Court Opinions. Indeed, a voluminous body of work has sought to explain how many times an opinion is cited \citep[e.g.,][]{cross2010determinants,benjamin2012standing}, when in a lifecycle an opinion is cited \citep[e.g.,][]{black2013citation,spriggs2001explaining}, and how many cases are cited by an opinion \citep[e.g.,][]{lupu2013strategic}---all focused on the US Supreme Court. One common feature of the research design in all of these studies is that the observations are at the case or case-time level. The outcome variables in these analyses are defined as measures of the number of citations to a case over a period of time, the number of citations to a case at a particular time, or a measurement on the cases cited by a case. None of these studies treats citation in its micro-level form---as a relationship between two opinions, the citing opinion and the cited opinion. We are aware of one prior study, \citet{clark2010locating}, in which a statistical model is used to analyze dyadic citations between cases. However, citet{clark2010locating} uses a dyadic latent variable model in order to estimate ideal points for Supreme Court Opinions, but does not use any explanatory variables to predict the formation of citation ties between opinions. We build upon this literature both methodologically and substantively. Methodologically, we develop a novel statistical framework for modeling directed dyadic citations. Second, we apply this methodology to a half-century of directed dyadic citations between US Supreme Court opinions.

There exist two broad reasons why empirical analyses of citations are best defined on the directed dyad level, not the case level. The first is that directed dyadic analyses can test both dyadic and case-level hypotheses. For example, case-level analyses can model whether opinions supported by a liberal majority coalition are more likely than those supported by a conservative majority coalition to be cited heavily in the future, but they cannot precisely model whether liberal cases will be cited more by liberal cases than by conservative cases. Thus, the first reason for analyzing citations at the dyadic level it to expand the set of hypotheses that can be represented in the model. The second reason for studying citations at the directed dyadic level is that, as articulated in the growing literature on legal citation networks, citations form complex networks in which a citation at one point in time may influence future citations. This phenomenon of complex dependence is very common in networks of many types, but processes specific to Supreme Court citations create interdependence in citations. For example, if opinion $i$ relies heavily on opinion $j$ as precedent, opinion $i$ is likely to discuss the legal basis for opinion $j$, and as a consequence cite some of the opinions cited by opinion $j$. Suppose opinion $k$ is cited by opinion $j$. Opinion $k$ is more likely to be cited by opinion $i$ because opinion $i$ relies heavily on $j$, and opinion $j$ cites $k$.  This is a special case of a very common process on networks referred to as ``triad closure''. Complex dependence is theoretically interesting on its own merits, but the effects of covariates cannot be reliably identified---either in terms of coefficient values or standard errors---without accounting for the interdependence inherent in networks \citep{desmaraisstatistical}.

In what follows we develop a theoretical case that citations on the US Supreme Court are characterized by forms of complex dependence that are common in networks. We then develop an extension of a model---the exponential random graph model---that can incorporate both exogenous covariates and complex forms of interdependence into a directed dyadic analysis of citations. Finally, we develop and estimate a specification of this model in an analysis of US Supreme Court citations between 1937 and 2001. We find robust support for the inherent complexity underpinning the formation of citation ties, and show that incorporating complex dependence into the model of citation formation significantly improves the model's predictive performance.

\section{Complex Interdependence in Supreme Court Citations}

When it comes to the development and testing of theory, the defining feature of networks is that the fundamental element under study---the relationship between two units (i.e., the citation from one opinion to another) is a piece of a complex web of relations that cannot be fully understood without considering how the singular relationship fits into that complex web.

@@

  \begin{figure}[H]
  \begin{tabular}{cc}
 (a) MQ Score Difference & (b) Ideological Breadth \\
\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mq_coef_trend.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/absdiffmq_coef_trend.pdf} \\

  (c) Same Issue & (d) Majority Size \\
\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/sameissue_coef_trend.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/numberjusticespro_coef_trend.pdf} \\

   (e) Reciprocity & (f) Transitivity \\
\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mutual_coef_trend.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/triangle_coef_trend.pdf} \\

    (g) Sociality & (h) Popularity \\
\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/o2star_coef_trend.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/i2star_coef_trend.pdf} \\

\end{tabular}
\caption{Effects}
 \label{fig:coeftrends}
\vspace{-.25cm}
\end{figure}
